---
title: "Exploratory Data Analysis"
output: html_notebook
---

This is a file for explorative data analysis of the premier league players.

```{r Import}
season17 <- read.csv("~/DSI-SRP1/season17.csv")
season18 <- read.csv("~/DSI-SRP1/season18.csv")
season19 <- read.csv("~/DSI-SRP1/season19.csv")
```

Checking out the datasets
```{r}
View(season17)
View(season18)
View(season19)
```

Loading tidyverse
```{r}
library(tidyverse)
```

Calculating correlation coefficients between goal scored and total fpl points
```{r}
message("Correlation between goal scored and total fpl points for 2016/17: ", with(season17, cor(goals_scored,total_points)))
message("Correlation between goal scored and total fpl points for 2017/18: ", with(season18, cor(goals_scored,total_points)))
message("Correlation between goal scored and total fpl points for 2018/19: ", with(season19, cor(goals_scored,total_points)))
```
Editing the datasets
```{r}
season17$position_index <- as_factor(season17$position_index)
season18$position_index <- as_factor(season18$position_index)
season19$position_index <- as_factor(season19$position_index)
```

Creating a new metric for calculation total fpl points to minutes played
```{r}
fpl_to_minute <- function(df) {
  df %>% 
    mutate(fpl_to_min = (total_points/minutes.played)*100)
}
season17 <- fpl_to_minute(season17)
season18 <- fpl_to_minute(season18)
season19 <- fpl_to_minute(season19)
```

Scatter plots graph to highlight how the goals scored vary with fpl points
```{r}
ggplot(season17, aes(goals_scored, total_points)) + geom_point(aes(color = position_index)) + labs(title = "Total FPL points vs Goal Scored in 2016/17")
ggplot(season18, aes(goals_scored, total_points)) + geom_point(aes(color = position_index)) + labs(title = "Total FPL points vs Goal Scored in 2017/18")
ggplot(season19, aes(goals_scored, total_points)) + geom_point(aes(color = position_index)) + labs(title = "Total FPL points vs Goal Scored in 2018/19")
```

Bar graphs to show how the number of players vary per team with regards to position
```{r}
ggplot(season17, aes(club_name)) + geom_bar(aes(fill = position_index),position = "fill") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Creating frequency plot for goal scored and how it varies for the players' position

```{r}
ggplot(season17, aes(x = goals_scored)) + geom_freqpoly(aes(color = position_index), binwidth = 2)
```
Creating boxplots to show the distiribution of the teams' fpl points according to their players
```{r}
ggplot(data = season17) + geom_boxplot(aes(x = club_name, y = total_points), na.rm = TRUE) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
write_csv(season17, "season17.csv")
write_csv(season18, "season18.csv")
write_csv(season19, "season19.csv")
```

Scatter plotsd showing the FPL points to minute ratio vs the total FPL points

```{r}
season17 %>% 
  filter(minutes.played >= 500) %>% 
  ggplot() + geom_point(aes(total_points, fpl_to_min, color = position_index), na.rm = TRUE) + labs(x = "Total FPL points", y = "FPL points-to-minute ratio", title = "FPL points to minutes played ratio vs Total FPL points for players\nwho have played more than 500 minutes")
``` 

```{r}
season18 %>% 
  filter(minutes.played >= 500) %>% 
  ggplot() + geom_point(aes(total_points, fpl_to_min, color = position_index), na.rm = TRUE) + labs(x = "Total FPL points", y = "FPL points-to-minute ratio", title = "FPL points to minutes played ratio vs Total FPL points for players\nwho have played more than 500 minutes")
```

```{r}
season19 %>% 
  filter(minutes.played >= 500) %>% 
  ggplot() + geom_point(aes(total_points, fpl_to_min, color = position_index), na.rm = TRUE) + labs(x = "Total FPL points", y = "FPL points-to-minute ratio", title = "FPL points to minutes played ratio vs Total FPL points for players\nwho have played more than 500 minutes")
```

Calculating the average fpl points per team

```{r}
avg_fpl_point <- function(df) {
  df %>% 
    group_by(club_name) %>%
    summarize(average = mean(total_points, na.rm = TRUE)) %>%
    ggplot() + geom_bar(aes(x = club_name, y = average, fill = club_name), stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(x = "Club Name", y = "Average FPL points", title = sprintf("The average FPL points achieved by each team in the season %s", unique(df$season)))
}
                                                                  
```

```{r, fig.width = 12, fig.height = 9}
avg_fpl_point(season17)
avg_fpl_point(season18)
avg_fpl_point(season19)
```

```{r}
goals_vs_assist <- function(df) {
  df %>%
    filter(minutes.played >= 500) %>%
    ggplot(aes(assists, goals_scored)) + geom_point(aes(color = position_index), position = "jitter") + labs(title = sprintf("Goals scored vs Assists during the %s season", unique(df$season)), color = "position index")
}
```

```{r, fig.height = 9, fig.width = 9}
goals_vs_assist(season17)
goals_vs_assist(season18)
goals_vs_assist(season19)
```


```{r}
fpl_point_rank <- function(df) {
  df %>% 
    mutate(ranking = dense_rank(desc(total_points))) %>%
    filter(ranking <= 10) %>%
    arrange(desc(total_points)) %>% 
    select(ranking, player_name, total_points) %>% 
    print() %>% 
    ggplot(aes(reorder(player_name, -total_points), total_points, fill = player_name)) + geom_bar(stat = "identity", color = "purple") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title = sprintf("The top 10 players with the highest FPL points in %s", unique(df$season)), x = "Player", y = "FPL points")
}
```

```{r, fig.height = 9, fig.width = 9}
fpl_point_rank(season17)
fpl_point_rank(season18)
fpl_point_rank(season19)
```

```{r}
library(modelr)
options(na.action = na.warn)
value <- function(df) {
  ggplot(df, aes(x = now_cost)) + geom_point(aes(y = total_points), color = "blue") + facet_wrap(~position_index)
}
```


```{r, fig.weight = 9, fig.height = 9}
value(season18)
value(season19)
```

